---
title: "Moran’s i"
author: "Samantha A. Alger"
date: "3/25/2017"
output: html_document
---
```{r}
ls()
rm(list=ls())

#Set Working Directory: 
setwd("~/AlgerProjects/2015_Bombus_Survey/CSV_Files")

#install package for Morans.i
install.packages(ape)

# get programs
library(ape)
library(plyr)
library(dplyr)


# read in data:
SurvData <- read.csv("MixedModelDF.csv", header=TRUE, sep = ",", stringsAsFactors=FALSE)

SpatialData <- read.csv("2015SurveyLocations.csv", header=TRUE, sep = ",", stringsAsFactors=FALSE)

head(SurvData)
head(SpatialData)

#merge the two datafiles by Site name
SpatialData <- merge(SurvData,SpatialData, by="site")

#Split by viruses
SpatDWV <- subset(SpatialData, target_name=="DWV")
SpatBQ <- subset(SpatialData, target_name=="BQCV")

#not normal! so log transform each one
#DWV:
hist(SpatDWV$BombPrev)
hist(SpatBQ$BombPrev)
SpatDWV$logPrev <- log(SpatDWV$BombPrev)

SpatDWV
hist(SpatDWV$logPrev)
SpatDWV$logPrev[SpatDWV$logPrev == "-Inf"] <- 0

SpatDWV
hist(SpatDWV$logPrev)

#BQCV
hist(SpatDWV$BombPrev)
hist(SpatBQ$BombPrev)
SpatBQ$logPrev <- (SpatBQ$BombPrev)^3

SpatBQ
hist(SpatBQ$logPrev)


```

#################################################################
Calculating Moran's I
#################################################################

Steps:
1. Generate a distance matrix

2. take inverse of the matrix values 
and replace the diagonal entries with zero

3. repeat for each virus

```{r}
#For DWV:
DWV.dists <- as.matrix(dist(cbind(SpatDWV$long, SpatDWV$lat)))

DWV.dists.inv <- 1/DWV.dists
diag(DWV.dists.inv) <- 0
 
DWV.dists.inv[1:5, 1:5]

#For BQCV:
BQ.dists <- as.matrix(dist(cbind(SpatBQ$long, SpatBQ$lat)))

BQ.dists.inv <- 1/BQ.dists
diag(BQ.dists.inv) <- 0
 
BQ.dists.inv[1:5, 1:5]

#Calculating Morans i:
Moran.I(SpatDWV$logPrev, DWV.dists.inv)
Moran.I(SpatBQ$logPrev, BQ.dists.inv)
Moran.I(SpatDWV$BombusViralLoad, DWV.dists.inv)
Moran.I(SpatBQ$BombusViralLoad, BQ.dists.inv)
Moran.I(SpatBQ$HBviralLoad, BQ.dists.inv)
Moran.I(SpatDWV$HBviralLoad, DWV.dists.inv)
